Work Fragmentation in Developer Interaction Data

### 2016-01-22 paper_2.tex
comments

\usepackage{graphics}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{theorem}
\usepackage{verbatim} % comentarios
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{mwe}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage[export]{adjustbox}

\usepackage{array}

@@

This work is an extension to the a published paper on 22nd IEEE International Conference on Software Analysis, Evolution, and Reengineering. Now, this improved version includes the same analysis previously made with a new dataset, the Eclipse Usage Data Collector (UDC) interaction data, which allows us to dig further into more specific aspects, resulting in the following two complementary research questions:
\begin{itemize}
\item \textbf{RQ4:} How is the developer's performance during the recovery time after an interruption? 
\item \textbf{RQ5:} What events are more common during recovery time?
\end{itemize}

@@

While these studies already show the effect of work fragmentation is hurtful, they are mostly observational studies over a limited time, and do not address specifically the sub-population of programmers.

@@

@@

@@

@@

\subsection{Selection}
We used the pre-processed version of the data that is published on Google BigQuery, by Murphy-Hill \cite{SnipesETALASD}. This is an alternative version of the original UDC dataset, which is cleaned and preprocessed, so that the transformation phase is simple and focused on our needs. Due to the magnitude of the dataset we only worked on a fragment of it. We took a random sample of 1,000 users that is larger than the dataset we were able to obtain for the Mylyn study.

@@

\subsection{Metrics Used in This Study}

Before introducing the results about the first research question, it is convenient to present the metrics used in this study.
We use the following six metrics to measure interruptions and productivity, while controlling for the unproductive time in a session, the length of time of the session itself, and the efficiency of the development that took place during the session.
\\

@@

@@

On the other hand, before the interruption there is a preparation phase when the interruption is imminent or expected. In this phase the programmer might leave notes or comments in the code to recover the context after the interruption \cite{PD10}. Also, when the interruption happened because of a problem found by the programmer, usually he tries to solve it by reading the code, using the debugger or reading comments in the control version tool; after this resources are depleted, the next option is to ask to teammates or other external resources \cite{LVD06}. The latest activities can represent an interruption of work.

@@

%We called \textit{recuperation time} the first intersect point between edition and navigation. In this sense, the  average recuperation time of an interruption occurs in the minute 12. 
\item \textit{On the left,} we observe that the number of events near the interruption also goes down well below the median value. This is might be because of two reasons: the programmer could have found a problem while coding, who at first will try to solve it by reading the code, switching out from the IDE, navigating the call stack and debugging, before going to ask another teammate according to LaToza et al. \cite{LVD06}; this set of actions end up with an interruption and reduce the observed activity within the IDE. Also, when the interruption is imminent or expected, the programmer make use of different suspension strategies like writing physical notes, making a mental note or leaving a reminder cue on the code or window, as mentioned by Parnin and DeLine \cite{PD10}. These activities reduce the activity before the gap and they are seen in the interaction data mostly as selection events. The latest can be better visualized in Figure \ref{fig:all_ints} with the UDC data as a significant decrease of edit events, being below the selection events approximately 10 minutes before the interruption.
%\item \textit{On the left,} we observe that the number of events near the interruption also goes down well below the median value. This is because work fragmentation is not only due to interruptions: programmers may for instance switch away from the IDE to look up information on the web or request it from other programmers. They are likely to do so because they are stuck, which would manifest itself by lower activity before the activity gap. We investigate this further below.

@@

We did so by computing Cohen's $d$ on the quantity of edits before and after the interruption. To obtain a significant effect, we need the presence of activity both before and after. Not all the interruptions meet this criteria however: some are located close to the start or the end of a session, or too close to another interruption.  In total, 53\% of the interruptions had 30 minutes before and after the interruption and were selected for the analysis.  The rest 47\% of the interruptions can not be used on this analysis due to the lack of time around them. %We used the Algorithm \ref{alg:local_effect} for the classification, and the 
Table \ref{tbl:local_effect} shows the applied thresholds and the results, accompanied with a typical example of an interruption in each category. This local analysis shows that there are indeed three well-defined groups of interruptions, with the two largest of them having clear effects on the activity in the session.

@@

We established the recovery time as the 15 minute period after an interruption, without considering the time consumed by the interruption itself. The length of the period was chosen according to the observations by Solingen et al. \cite{SBV98}.

@@

The duration of the peak varies among the also called working spheres, as mentioned by Gonzalez and Mark \cite{MGH05}. On average, the peak lasts for 12 minutes, while the maximum registered is of 145 minutes. The distribution of duration of the peaks is reported in Table \ref{tbl:dist_peaks}.

@@

Observing the user actvivity throughout a working session might be the best way to understand the effect of events that occur on the working place, like work fragmentation. However, implementing a study of this kind can be very expensive, for the time and material need. For example, to have the same amount of information we have with UDC it would require to monitor the user activity for a year or two and have a record of all the activities performed, like the number and duration of the interruptions. Moreover, this kind of research require on-site studies, exhaustive monitoring and interviews, which in some cases could be expensive or difficult to perform.

@@

Another threat comes from the metrics we use. Our choices are limited by the data recorded by Mylyn. We believe our three indicator (edits per minute, selections per minute, and edit ratio) provide a reasonable measure of productivity (especially with the addition of edit ratio, used in other work \cite{KM06}). However we can not exclude that other indicators of productivity (such as actual changes~\cite{RL10}) may yield different results.

>>>>

\usepackage{graphics}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{theorem}
\usepackage{verbatim} % comentarios
\usepackage{amsmath}
%\usepackage{algorithm}
%\usepackage{algpseudocode}
\usepackage{mwe}
\usepackage{hyperref}
\usepackage{xspace}
\usepackage[export]{adjustbox}

\usepackage{array}
\newcommand\RR[1]{\textbf{Romain}{#1}}

@@

This work is an extension to a published paper in the 22nd IEEE International Conference on Software Analysis, Evolution, and Reengineering. This improved version includes the same analysis previously made with a new dataset, the Eclipse Usage Data Collector (UDC) interaction data, which allows us to dig further into more specific aspects, resulting in the following two complementary research questions:
\begin{itemize}
\item \textbf{RQ4:} How is the developer's performance during the recovery time after an interruption? 
\item \textbf{RQ5:} What events are more common during recovery time?
\end{itemize}

\RR{For each research question, add a sentence explaining why we look into it}

@@

While these studies already show the effect of work fragmentation is hurtful, they are mostly observational studies over a limited time, and do not address specifically the sub-population of programmers.
\RR{Some rationale as to why this population is specific enough to be studied in isolation. For instance, that programmers have to maintain very complicated mental contexts (e.g. understanding complex programs). Perhaps Parnin has something about it (Memery failure paper?)}

\RR{Make sure the new papers asked by Reviewer 1 are also mentioned in the related work}

@@

\RR{Perhaps another example here: it was not clear for one of the reviewer why this was necessary. E.g.: if I scroll down in the outline view with the keyboard I will be generating many selection events. Check other characteristics too (agregate, long).}

@@

\RR{motivate more the time series stuff}

@@

\RR{Add emphasis to the paragraph above since reviewers missed it (asking for motivation for the thresholds)}

@@

\subsection{Selection}
We used the pre-processed version of the data that is published on Google BigQuery, by Murphy-Hill et al. \cite{SnipesETALASD}. This is an alternative version of the original UDC dataset, which is cleaned and preprocessed, so that the transformation phase is simple and focused on our needs. Due to the magnitude of the dataset we only worked on a fragment of it. We took a random sample of 1,000 users that is larger than the dataset we were able to obtain for the Mylyn study.

@@

\subsection{Metrics Used in This Study}
\RR{consider moving subsection out of RQ1}
Before introducing the results about the first research question, it is convenient to present the metrics used in this study.
We use the following six metrics to measure interruptions and productivity, while controlling for the unproductive time in a session, the length of time of the session itself, and the efficiency of the development that took place during the session.
\\

@@

\RR{motivate more edit ratio, and add here the correlation analysis}

@@

\RR{expand this paragraph, since a reviewer asks explicitely for that}
On the other hand, before the interruption there is a preparation phase when the interruption is imminent or expected. In this phase the programmer might leave notes or comments in the code to recover the context after the interruption \cite{PD10}. Also, when the interruption happened because of a problem found by the programmer, usually he tries to solve it by reading the code, using the debugger or reading comments in the control version tool; after this resources are depleted, the next option is to ask to teammates or other external resources \cite{LVD06}. The latest activities can represent an interruption of work.

@@

%We called \textit{recuperation time} the first intersect point between edition and navigation. In this sense, the  average recuperation time of an interruption occurs in the minute 12. 
\item \textit{On the left,} we observe that the number of events near the interruption also goes down well below the median value. This might be because of two reasons: the programmer could have found a problem while coding, who at first will try to solve it by reading the code, switching out from the IDE, navigating the call stack and debugging, before going to ask another teammate according to LaToza et al. \cite{LVD06}; this set of actions end up with an interruption and reduce the observed activity within the IDE. Also, when the interruption is imminent or expected, the programmer make use of different suspension strategies like writing physical notes, making a mental note or leaving a reminder cue on the code or window, as mentioned by Parnin and DeLine \cite{PD10}. These activities reduce the activity before the gap and they are seen in the interaction data mostly as selection events. The latest can be better visualized in Figure \ref{fig:all_ints} with the UDC data as a significant decrease of edit events, being below the selection events approximately 10 minutes before the interruption.
%\item \textit{On the left,} we observe that the number of events near the interruption also goes down well below the median value. This is because work fragmentation is not only due to interruptions: programmers may for instance switch away from the IDE to look up information on the web or request it from other programmers. They are likely to do so because they are stuck, which would manifest itself by lower activity before the activity gap. We investigate this further below.

@@

We did so by computing Cohen's $d$ on the quantity of edits before and after the interruption. To obtain a significant effect, we need the presence of activity both before and after. Not all the interruptions meet this criteria however: some are located close to the start or the end of a session, or too close to another interruption.  In total, 53\% of the interruptions had 30 minutes before and after the interruption and were selected for the analysis.  The remaining 47\% of the interruptions can not be used on this analysis due to the lack of time around them. %We used the Algorithm \ref{alg:local_effect} for the classification, and the 
Table \ref{tbl:local_effect} shows the applied thresholds and the results, accompanied with a typical example of an interruption in each category. This local analysis shows that there are indeed three well-defined groups of interruptions, with the two largest of them having clear effects on the activity in the session.

@@

We established the recovery time as the 15 minute period after an interruption, without considering the time consumed by the interruption itself. The length of the period was chosen according to the observations by Solingen et al. \cite{SBV98}.
\RR{Add details of that paper to better justify}

@@

The duration of the peak varies among the also called working spheres \RR{add details}, as mentioned by Gonzalez and Mark \cite{MGH05}. On average, the peak lasts for 12 minutes, while the maximum registered is of 145 minutes. The distribution of duration of the peaks is reported in Table \ref{tbl:dist_peaks}.

@@

Observing the user activity throughout a working session might be the best way to understand the effect of events that occur on the working place, like work fragmentation. However, implementing a study of this kind can be very expensive, for the time and material need. For example, to have the same amount of information we have with UDC it would require to monitor the user activity for a year or two and have a record of all the activities performed, like the number and duration of the interruptions. Moreover, this kind of research require on-site studies, exhaustive monitoring and interviews, which in some cases could be expensive or difficult to perform.

@@

Another threat comes from the metrics we use. Our choices are limited by the data recorded by Mylyn. We believe our three indicator (edits per minute, selections per minute, and edit ratio) provide a reasonable measure of productivity (especially with the addition of edit ratio, used in other work \cite{KM06}). However we can not exclude that other indicators of productivity (such as actual changes~\cite{RL10}) may yield different results. 
\RR{add a bit stuff about edit ratio to keep one reviewer happy}
